---
title: "Analysis report of melanoma scRNA-Seq data"
author: "Maria Schmidt"
affiliation: "IZBI, Interdisciplinary Centre for Bioinformatics, Leipzig University, Germany"
date: today
format:
  html:
    toc: true
    number-sections: true
    code-fold: true
  pptx: default
  pdf: default
execute: 
  echo: true
  error: false
  warning: false
link-citations: true
bibliography: "library.bib"
embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r libraries and load data}
pacman::p_load(Seurat, tidyverse, here, ggplot2, pals, patchwork, liana)
source(here("src","seurat-functions.r"))
source(here("src","liana-functions.r"))
source(here("src","paths.r"))

filename <- file.path(output_data_path, "integrated-seurat-obj.RData")
load(filename)
filename <- file.path(output_data_path, "color-list.RData")
load(filename)   
seurat_colors <- scales::hue_pal()(length(table(obj_integr@meta.data[, "seurat_clusters"])))
# seurat_cluster_data <- data.frame(
#     table(obj_integr@meta.data[, "seurat_clusters"]),
#     "color" = seurat_colors
# )
celltype_colors <- color_list[["predicted.id.Cell_type_nicknames"]]
# celltype_cluster_data <- data.frame(
#     table(obj_integr@meta.data[, "predicted.id.Cell_type_nicknames"]),
#     "color" = celltype_colors
# )
# filename <- here(output_data_path, "predicted.id.Cell_type_nicknames_marker.RData")
# load(filename)
filename <- here(output_data_path, "reynolds-skin-marker.RData")
load(filename) 
filename <- here(output_data_path, "integrated-liana-obj-Cell_type_nicknames_nevus-marked.RData")
load(filename)
filename <- here(literature_path, "tsoi.marker.list.RData")
load(filename)
filename <- here(literature_path, "belote.list.RData")
load(filename) 

```

# LIANA-analysis

## communication frequency

The plot visualizes the frequencies of significant interactions (p <= 0.01) for each pair of potentially communicating cell types from sender to receiver.

```{r liana_freq, echo=FALSE, fig.width=20, fig.height=20}

liana_obj %>%
  filter(aggregate_rank <= 0.01) %>%
  heat_freq()

```


```{r , echo=FALSE, fig.width=20, fig.height=20}
dendritic <- grep("DC", unique(obj_integr@meta.data[,"predicted.id.Cell_type_nicknames_mels_tsoi-classes_nevus_marked"]), value = TRUE)
dendritic <- dendritic[-grep("nev", dendritic)]
tcells <- c("Tc", "Th", "Treg")
nkcells <- grep("NK", unique(obj_integr@meta.data[,"predicted.id.Cell_type_nicknames_mels_tsoi-classes_nevus_marked"]), value = TRUE)
lymphocytes <- c(tcells, nkcells, "Plasma cell" )
macrophages <- grep("Mac", unique(obj_integr@meta.data[,"predicted.id.Cell_type_nicknames_mels_tsoi-classes_nevus_marked"]), value = TRUE)
macrophages <- macrophages[-grep("nev", macrophages)]
langerhans <- grep("LC", unique(obj_integr@meta.data[,"predicted.id.Cell_type_nicknames_mels_tsoi-classes_nevus_marked"]), value = TRUE)
langerhans <- langerhans[-grep("nev", langerhans)]
inflammatory <- sort(c(dendritic, tcells, macrophages, langerhans, "Mast cell", "Monocyte", "Plasma cell" ))
mast <-  grep("Mast", unique(obj_integr@meta.data[,"predicted.id.Cell_type_nicknames_mels_tsoi-classes_nevus_marked"]), value = TRUE)
mast <- mast[-grep("nev", mast)]
melanocytes <- grep("Mel_", unique(obj_integr@meta.data[,"predicted.id.Cell_type_nicknames_mels_tsoi-classes_nevus_marked"]), value = TRUE)
melanocytes <- melanocytes[-grep("nev", melanocytes)]
mel_and_mast <- unique(c(mast, melanocytes))
```

## t cells -> melanocytes

All significant immune-melanoma interactions with p <= 0.01. Sender region is on the left and receiver region on the right. For more information on LIANA see Dimitrov et al. 2022

```{r , echo=FALSE, fig.width=20, fig.height=10}

liana_subset <- liana_obj %>%
  filter(aggregate_rank < 0.01) %>%
  filter(source %in%  tcells) %>%
  filter(target %in%  melanocytes) 
  
  
  if (nrow(liana_subset) > 0) {
    my_liana_dotplot2(liana_subset,
                      source_groups = tcells,
                      target_groups = melanocytes) %>% print()
  }
```

## t cells -> melanocytes + mast cells

```{r , echo=FALSE, fig.width=20, fig.height=10}

liana_subset <- liana_obj %>%
  filter(aggregate_rank < 0.01) %>%
  filter(source %in%  tcells) %>%
  filter(target %in%  mel_and_mast) 
  
  
  if (nrow(liana_subset) > 0) {
    my_liana_dotplot2(liana_subset,
                      source_groups = tcells,
                      target_groups = mel_and_mast) %>% print()
  }
```

## melanocytes -> t cells

```{r , echo=FALSE, fig.width=25, fig.height=15}

liana_subset <- liana_obj %>%
  filter(aggregate_rank < 0.01) %>%
  filter(source %in%  melanocytes) %>%
  filter(target %in%  tcells) 
  
  
  if (nrow(liana_subset) > 0) {
    my_liana_dotplot2(liana_subset,
                      source_groups = melanocytes,
                      target_groups = tcells) %>% print()
  }
```

## melanocytes + mast cells -> t cells

```{r , echo=FALSE, fig.width=25, fig.height=35}

liana_subset <- liana_obj %>%
  filter(aggregate_rank < 0.01) %>%
  filter(source %in%  mel_and_mast) %>%
  filter(target %in%  tcells) 
  
  
  if (nrow(liana_subset) > 0) {
    my_liana_dotplot2(liana_subset,
                      source_groups = mel_and_mast,
                      target_groups = tcells) %>% print()
  }
```






## dendritic cells -> melanocytes

```{r , echo=FALSE, fig.width=30, fig.height=20}

liana_subset <- liana_obj %>%
  filter(aggregate_rank < 0.01) %>%
  filter(source %in%  dendritic) %>%
  filter(target %in%  melanocytes) 
  
  
  if (nrow(liana_subset) > 0) {
    my_liana_dotplot2(liana_subset,
                      source_groups = dendritic,
                      target_groups = melanocytes) %>% print()
  }
```

## dendritic cells -> melanocytes + mast cells

```{r , echo=FALSE, fig.width=30, fig.height=20}

liana_subset <- liana_obj %>%
  filter(aggregate_rank < 0.01) %>%
  filter(source %in%  dendritic) %>%
  filter(target %in%  mel_and_mast) 
  
  
  if (nrow(liana_subset) > 0) {
    my_liana_dotplot2(liana_subset,
                      source_groups = dendritic,
                      target_groups = mel_and_mast) %>% print()
  }
```

## melanocytes -> dendritic cells

```{r , echo=FALSE, fig.width=30, fig.height=20}

liana_subset <- liana_obj %>%
  filter(aggregate_rank < 0.01) %>%
  filter(source %in%  melanocytes) %>%
  filter(target %in%  dendritic) 
  
  
  if (nrow(liana_subset) > 0) {
    my_liana_dotplot2(liana_subset,
                      source_groups = melanocytes,
                      target_groups = dendritic) %>% print()
  }
```

## melanocytes + mast cells -> dendritic cells

```{r , echo=FALSE, fig.width=30, fig.height=35}

liana_subset <- liana_obj %>%
  filter(aggregate_rank < 0.01) %>%
  filter(source %in%  mel_and_mast) %>%
  filter(target %in%  dendritic) 
  
  
  if (nrow(liana_subset) > 0) {
    my_liana_dotplot2(liana_subset,
                      source_groups = mel_and_mast,
                      target_groups = dendritic) %>% print()
  }
```



## dendritic cells -> lymphocytes

p <= 0.001

```{r , echo=FALSE, fig.width=30, fig.height=20}

liana_subset <- liana_obj %>%
  filter(aggregate_rank < 0.001) %>%
  filter(source %in%  dendritic) %>%
  filter(target %in%  lymphocytes) 
  
  
  if (nrow(liana_subset) > 0) {
    my_liana_dotplot2(liana_subset,
                      source_groups = dendritic,
                      target_groups = lymphocytes) %>% print()
  }
```


## lymphocytes -> dendritic cells

```{r , echo=FALSE, fig.width=25, fig.height=20}

liana_subset <- liana_obj %>%
  filter(aggregate_rank < 0.01) %>%
  filter(source %in%  lymphocytes) %>%
  filter(target %in%  dendritic) 
  
  
  if (nrow(liana_subset) > 0) {
    my_liana_dotplot2(liana_subset,
                      source_groups = lymphocytes,
                      target_groups = dendritic) %>% print()
  }
```

## natural killer cells -> melanocytes

p <= 0.05

```{r , echo=FALSE, fig.width=20, fig.height=15}

liana_subset <- liana_obj %>%
  filter(aggregate_rank < 0.05) %>%
  filter(source %in%  nkcells) %>%
  filter(target %in%  melanocytes) 
  
  
  if (nrow(liana_subset) > 0) {
    my_liana_dotplot2(liana_subset,
                      source_groups = nkcells,
                      target_groups = melanocytes) %>% print()
  }
```


## melanocytes -> natural killer cells

```{r , echo=FALSE, fig.width=25, fig.height=20}

liana_subset <- liana_obj %>%
  filter(aggregate_rank < 0.05) %>%
  filter(source %in%  melanocytes) %>%
  filter(target %in%  nkcells) 
  
  
  if (nrow(liana_subset) > 0) {
    my_liana_dotplot2(liana_subset,
                      source_groups = melanocytes,
                      target_groups = nkcells) %>% print()
  }
```

## natural killer cells -> melanocytes + mast cells

p <= 0.05

```{r , echo=FALSE, fig.width=20, fig.height=15}

liana_subset <- liana_obj %>%
  filter(aggregate_rank < 0.05) %>%
  filter(source %in%  nkcells) %>%
  filter(target %in%  mel_and_mast) 
  
  
  if (nrow(liana_subset) > 0) {
    my_liana_dotplot2(liana_subset,
                      source_groups = nkcells,
                      target_groups = mel_and_mast) %>% print()
  }
```


## melanocytes + mast cells -> natural killer cells

```{r , echo=FALSE, fig.width=25, fig.height=20}

liana_subset <- liana_obj %>%
  filter(aggregate_rank < 0.05) %>%
  filter(source %in%  mel_and_mast) %>%
  filter(target %in%  nkcells) 
  
  
  if (nrow(liana_subset) > 0) {
    my_liana_dotplot2(liana_subset,
                      source_groups = mel_and_mast,
                      target_groups = nkcells) %>% print()
  }
```